Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-08-15 manuscript/Appendix2.Rmd
updates all model related stuff to ARK specification

\begin{center}
\textbf{\Large{Appendix 2}} \\
A.T. Tredennick, A.R. Kleinhesselink, \& P.B. Adler \\
``Ecosystem and community resistance...'' \\
\emph{PeerJ}
\end{center}

\section{Random Slopes, Random Intercepts Model}
\subsection{Model Description}
Our hierarchical Bayesian model with random intercepts and random slopes allowed us to test for differences among treatments in the relationship between ANPP and growing season precipitation, and allowed us to account for the non-independence of observations through time within a plot.
The random effects structure mirrors the nested structure of the data: plot-level parameters are nested within treatment-level parameters, which are nested within overall parameters.
In what follows, $\textbf{X}$ is the design matrix including a column of 1s for intercepts and a column of precipitation values for each observation, $\textbf{y}$ is a row vector of log(ANPP) values for each observation, *i* references observations, *j* references plots, and *k* references treatment.
The notation *i(j(k))* reads as 'observation *i* associated with plot *j* associated with treatment *k*'.
Recall we have three treatments: drought (~50\% decrease in precipitation), control (ambient precipitation), and irrigation (~50\% increase in precipitation).

We assume the observations are conditionally Gaussian,

@@

\begin{align}
y_{i(j(k))} \sim \text{Normal}\left(\mu_{i(j(k))}, \sigma^2_k \right),
\end{align}

\noindent{}where $\mu_{i(j(k))}$ is the determinstic expectation from the regression model,

@@

\begin{align}
\mu_{i(j(k))} = \textbf{x}_i'\boldsymbol{\beta_{j(k)}}.
\end{align}

Regression parameters (the intercept and slope in the parameter vector $\boldsymbol{\beta}$) were modeled hierarchically,

@@

\begin{align}
\boldsymbol{\beta}_{j(k)} &\sim \text{MVN} \left( \boldsymbol{\beta}_k, \boldsymbol{\Sigma}(k) \right), \\
\boldsymbol{\beta}_{k} &\sim \text{MVN} \left( \boldsymbol{\beta}, \boldsymbol{\Sigma}  \right), \\
\boldsymbol{\beta} &\sim \text{Normal} \left( 0, 1 \right),
\end{align}

\noindent{} where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\boldsymbol{\Sigma}(k)$).
The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.

We fit the model using MCMC as implemented in the software Stan [@stan2016].
Our Stan code is below.
All code necessary to reproduce our results is in the supplementary code set (Appendix 3).

\subsection{Stan Code}
```{r stan1, echo=T, eval=F}
data {
  int<lower=0> Npreds;        # number of covariates, including intercept
  int<lower=0> Nplots;        # number of plots
  int<lower=0> Ntreats;       # number of treatments
  int<lower=0> Nobs;          # number of observations
  int<lower=0> Nppts;         # Number of precip levels to predict
  vector[Nobs] y;             # vector of observations
  row_vector[Npreds] x[Nobs]; # design matrix
  matrix[Nppts,Npreds] newx;  # design matrix for predictions 
  matrix[Npreds,Npreds] R;	  # priors for covariance matrix
  int plot_id[Nobs];          # vector of plot ids
  int treat_id[Nobs];         # vector of treatment ids
}

parameters {
  vector[Npreds] beta_plot[Nplots]; # a unique vector matrix for each plot
  vector[Npreds] beta_treat[Ntreats]; # a unique vector matrix for each treatment
	vector[Npreds] beta_mu; # overall coefficients
	cov_matrix[Npreds] Sigma; # covariance matrix for treatment-level coefficients
	cov_matrix[Npreds] Sigma_plot[Ntreats]; # unique covariance matrix for plot-level coefs by treatment
	vector<lower=0>[Ntreats] sd_y; # treatment-level observation std. dev.
}

transformed parameters {
  vector[Nobs] yhat; # vector of expected values (predictions)
  for (i in 1:Nobs)
    # regression model for expected values (one for each plot-year)
    yhat[i] = x[i]*beta_plot[plot_id[i]]; 
}

model {
  ####  PRIORS
  for(i in 1:Nplots)
    # plot-level coefficients vary normally around treatment coefs
		beta_plot[i] ~ multi_normal(beta_treat[treat_id[i]], Sigma_plot[treat_id[i]]);

	for(i in 1:Ntreats){
	  # treatment-level coefficients vary normally around overall coefficients
	  beta_treat[i] ~ multi_normal(beta_mu, Sigma);

	  # priors on covariance of effects at plot-level
	  Sigma_plot[i] ~ inv_wishart(Npreds+1, R);
	}

	beta_mu ~ normal(0,1); # priors on overall effects
	Sigma ~ inv_wishart(Npreds+1, R); # priors on covariance of effects at treatment-level
	sd_y ~ weibull(2,1); # priors on observation std. dev. for each treatment

	####  LIKELIHOOD
	for(i in 1:Nobs)
	  # observations vary normally around expected values
	  y[i] ~ normal(yhat[i], sd_y[treat_id[i]]);
}

generated quantities {
  vector[Nppts] ypreds[Ntreats];
  vector[Nppts] ydiff_control_drought;
  vector[Nppts] ydiff_control_irrigate;
  vector[2] inter_diffs;
  for(i in 1:Ntreats)
    ypreds[i] = newx*beta_treat[i]; # mean predictions for each treatment

  # difference between mean predictions
  ydiff_control_drought = ypreds[1] - ypreds[2]; 
  ydiff_control_irrigate = ypreds[1] - ypreds[3];

  # difference between control and treatment in avg ppt year
  inter_diffs[1] = beta_treat[1][1] - beta_treat[2][1]; # control - drought
  inter_diffs[2] = beta_treat[1][1] - beta_treat[3][1]; # control - irrigation
}

@@

>>>>

\begin{center}
\textbf{\Large{Appendix 2}} \\
A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
\emph{PeerJ}
\end{center}

\section{Random Slopes, Random Intercepts Model}
\subsection{Model Description}
Our hierarchical Bayesian model allows us to test for differences among treatments in the relationship between ANPP and soil moisture, and allowed us to account for the non-independence of observations through time within a plot.
Treatment differences are modeled as fixed effects, which are modified by plot-level random effects.
In what follows, $\textbf{X}$ is the fixed effects design matrix including:

1. a column of 1s for intercepts
2. a column of continuous values for the scaled volumetric water content for each observation
3. a column of 0s or 1s indicating whether the observation is from a drought treatment 
4. a column of 0s or 1s indicating whether the observation is from an irrigation treatment
5. a column of continuous values for the interaction between scaled volumetric water content and the drought treatment indicator (column 2)
6. a column of continuous values for the interaction between scaled volumetric water content and the irrigation treatment indicator (column 3)

The first seven rows of the fixed effects design matrix (six control plots and one drought plot) is as follows:

```{r fixed_design, echo=F, eval=T, message=F, warning=F}
library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots
source("../code/read_format_data.R") # load data
lmod <- lm(log(anpp) ~ vwc_scaled*Treatment, anpp_data)
x    <- model.matrix(lmod) 
colnames(x) <- c("Int", "VWC", "Drought", "Irrig", "Drought:VWC", "Irrig:VWC")
head(x, 7)
```

\noindent{}Note that within a year, all plots within a treatment share the same value of volumetric water content. This is because we could not monitor soil moisture in each plot, and instead used sparse observations to model average soil moisture in each treatment in each year (see main text). Using this design matrix, we can estimate six fixed effects ($\boldsymbol{\beta}$s):

1. the intercept of the soil moisture-ANPP relationship for control plots
2. the slope of the soil moisture-ANPP relationship for control plots
3. the intercept offset for drought plots
4. the intercept offset for irrigation plots
5. the slope offset for drought plots
6. the slope offset for irrigation plots

\noindent{}We are particularly interested in the slope offsets, because these allow us to test whether the slopes for drought or irrigation are different from the control slope. If the slope offset is different from zero, this indicates the slopes are different. We assess whether the slope offsets are different from zero by calculating the probability that the posterior distribution is less or greater than zero (one-tailed tests). If the probability is greater than 0.95, then there is strong evidence that the slope offset is different from zero, and thus different from the control treatment slope.

To account for the fact that observations within plots through time are not independent, we include random effects that modify the fixed effects in each plot. We model these random effects ($\boldsymbol{\gamma}$s) as offsets drawn from a multivariate normal distribution with mean 0 and a variance-covariance matrix ($\Sigma$) that includes a covariance between the intercept and slope offsets. We implement this random effects structure by including a random effects design matrix ($\textbf{Z}$) with a column of 1s for intercept offsets and a column of continuous values for the volumetric water content for each observation.

Lastly, to account for unkown variation across years, we include random year effects. These year effects ($\boldsymbol{\eta}$s) act as offsets on the intercept.

Putting it all together, our model is defined mathematically as follows, where *i* denotes observation, *j* denotes plot, and *t* denotes year.
We assume the observations are conditionally Gaussian,

@@

\begin{align}
\textbf{y} &\sim \text{Normal} \left(\boldsymbol{\mu}, \sigma^2 \right),
\end{align}

\noindent{}where $\boldsymbol{\mu}$ is the determinstic expectation from the regression model,

@@

\begin{align}
\mu_{i} &= \boldsymbol{\beta}\textbf{x}_i + \boldsymbol{\gamma}_{j(i)}\textbf{z}_i + \eta_t.
\end{align}

\noindent{}All fixed effect $\beta$s were drawn from normally distributed priors with mean 0 and standard deviation 1: $\boldsymbol{\beta} \sim \text{Normal}(0,1)$. $\gamma$ random effects were drawn from a multivariate prior centered on zero with a shared variance covariance matrix:

@@

\begin{align}
\boldsymbol{\gamma} \sim \text{MVN} \left( 0, \Sigma \right), \\
\Sigma \sim \text{Cholesky...}
\end{align}

\noindent{The random year effects ($\boldsymbol{\eta}$) are modeled as intercept offsets centered on zero with a shared variance ($\sigma_{yr}$): $\boldsymbol{\gamma} \sim \text{Normal} \left( 0, \sigma_{yr}  \right)$.

We fit the model using MCMC as implemented in the software Stan [@stan2016].
Our Stan code is below.
All code necessary to reproduce our results has been archived on Figshare (*link here*) and released on GitHub (https://github.com/atredennick/usses_water/releases/v0.1).

\newpage{}

\subsection{Stan Code}
```{r stan1, echo=T, eval=F}
data {
  int<lower=0> Npreds;         # number of covariates, including intercept
  int<lower=0> Npreds2;        # number of random effect covariates
  int<lower=0> Nplots;         # number of plots
  int<lower=0> Ntreats;        # number of treatments
  int<lower=0> Nobs;           # number of observations
  int<lower=0> Nppts;          # number of precip levels to predict
  int<lower=0> Nyears;         # number of years
  vector[Nobs] y;              # vector of observations
  row_vector[Npreds] x[Nobs];	 # design matrix for fixed effects
  row_vector[Npreds2] z[Nobs]; # simple design matrix for random effects 
  matrix[Nppts,Npreds] newx;   # design matrix for predictions 
  matrix[Npreds2,Npreds2] R;	 # priors for covariance matrix
  int plot_id[Nobs];           # vector of plot ids
  int treat_id[Nobs];          # vector of treatment ids
  int year_id[Nobs];           # vector of year ids
}

parameters {
	vector[Npreds] beta;	             # overall coefficients
	vector[Nyears] year_off;           # vector of year effects
	cholesky_factor_corr[Npreds2] L_u; # cholesky factor of plot random effect corr matrix
  vector[Npreds2] beta_plot[Nplots]; # plot level random effects
	vector<lower=0>[Npreds2] sigma_u;  # plot random effect std. dev.
	real<lower=0> sd_y;                # treatment-level observation std. dev.
	real<lower=0> sigma_year;          # year std. dev. hyperprior
}

transformed parameters {
  vector[Nobs] yhat;               # vector of expected values
  vector[Npreds2] u[Nplots];       # transformed plot random effects
  matrix[Npreds2,Npreds2] Sigma_u; # plot ranef cov matrix

  Sigma_u = diag_pre_multiply(sigma_u, L_u); # plot-level covariance matrix 
  for(j in 1:Nplots)
    u[j] = Sigma_u * beta_plot[j]; # plot random intercepts and slopes

  # regression model for expected values (one for each plot-year)
  for (i in 1:Nobs)
    yhat[i] = x[i]*beta + z[i]*u[plot_id[i]] + year_off[year_id[i]]; 
}

model {
  ####  PRIORS
  sigma_year ~ cauchy(0,2)
  year_off ~ normal(0,sigma_year); # priors on year effects, shared variance
  beta ~ normal(0,1);	             # priors on treatment coefficients
  L_u ~ lkj_corr_cholesky(1);      # prior on the cholesky factor which controls the 
                                   # correlation between plot level treatment effects

  for(i in 1:Nplots)
		beta_plot[i] ~ normal(0,1); # plot-level coefficients for intercept and slope

	####  LIKELIHOOD
  for(i in 1:Nobs)
    y[i] ~ normal(yhat[i], sd_y); # observations vary normally around expected values
}
```

\newpage{}

# Figures

```{r scatterplot, fig.cap="Scatterplot, on the arithmetic scale, of the data and model estimates shown as solid lines. Model estimates come from treatment level coefficients (colored lines).", echo=FALSE, message=FALSE}
library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots

source("../code/read_format_data.R") # load data
all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")

param_labels <- c("Control::Intercept",
                  "Drought::Intercept", 
                  "Irrigation::Intercept",
                  "Control::Slope", 
                  "Drought::Slope", 
                  "Irrigation::Slope")

treat_effects <- data.frame(extract(all_fit, pars = 'beta')) %>% 
  mutate(iteration = row_number()) %>% 
  gather(param_name, estimate, starts_with('beta')) %>%
  mutate(param_name = factor(param_name, labels = param_labels)) %>% 
  spread(param_name, estimate) %>% 
  mutate(`Drought::Intercept` = `Drought::Intercept` + `Control::Intercept`, 
         `Irrigation::Intercept` = `Irrigation::Intercept` + `Control::Intercept`,  
         `Drought::Slope` = `Drought::Slope` + `Control::Slope`, 
         `Irrigation::Slope` = `Irrigation::Slope` + `Control::Slope`) %>%
  gather(param_name, estimate, -iteration)

mean_effects <- treat_effects %>%
  group_by(param_name) %>%
  summarise(mean_estimate = mean(estimate))

regression <- mean_effects %>%
  separate(param_name, into = c("Treatment","Coefficient"), sep = "::") %>%
  spread(Coefficient, mean_estimate)

lmod <- lm(log(anpp) ~ vwc_scaled, anpp_data)
x    <- model.matrix(lmod) 
newx <- data.frame(x = seq(min(x[,2]), max(x[,2]), length.out = 50)) %>%
  mutate(id = paste("x",1:50,sep = "::")) %>%
  spread(id, x)
newx <- rbind(rbind(newx,newx),newx)
newx$Treatment <- regression$Treatment

regression <- regression %>%
  left_join(newx, by="Treatment") %>%
  gather(xid, newx, -Treatment, -Intercept, -Slope)
regression$estimate <- with(regression, Intercept+Slope*newx)

mean_log_anpp <- mean(log(anpp_data$anpp))
sd_log_anpp <- sd(log(anpp_data$anpp))
mean_vwc <- mean(anpp_data$total_seasonal_vwc)
sd_vwc <- sd(anpp_data$total_seasonal_vwc)

regression <- regression %>%
  mutate(backtrans_estimate = estimate*sd_log_anpp+mean_log_anpp,
         backtrans_vwc = newx*sd_vwc+mean_vwc)

ggplot(anpp_data, aes(x=total_seasonal_vwc,y=anpp))+
  geom_point(shape=21,color="grey25",alpha=0.8,aes(fill=Treatment))+
  geom_line(data=regression, aes(x=backtrans_vwc, y=exp(backtrans_estimate), color=Treatment), size=1)+
  scale_fill_brewer(palette = "Set2")+
  scale_color_brewer(palette = "Set2")+
  xlab("March-June Cumulative VWC")+
  ylab(expression(paste("ANPP (g ",m^2,")")))+
  theme_bw()+
  theme(panel.grid.minor = element_blank())

@@

\newpage{}

```{r year_effects, fig.cap="Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.", echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)    # Data munging
library(dplyr)        # Data summarizing
library(ggthemes)     # Pleasing ggplot themes
library(ggjoy)        # Joy plots!
library(stringr)      # Working with strings
library(rstan)        # For MCMC and Stan objects
library(gridExtra)    # For combining ggplot objects
library(viridis)      # Pleasing color palette
library(RColorBrewer) # More color palettes
library(cowplot)      # For combining ggplots

all_fit <- readRDS("../results/randcoefs_alltreatments_fit.RDS")

year_id_names <- data.frame(year_id = c(1:5),
                            year_name = c(1:5))
betas <- reshape2::melt(rstan::extract(all_fit, pars="year_off")) %>%
  rename(iteration = iterations, year_id = Var2, estimate = value, stan_name = L1) %>%
  left_join(year_id_names, by="year_id")

ggplot(betas, aes(x=estimate, y=as.factor(year_id), fill=as.factor(year_id), color=as.factor(year_id), height = ..density..))+
  geom_vline(aes(xintercept=0), linetype=2, color="grey45")+
  geom_joy(stat="density", adjust=3, alpha=0.9)+
  scale_x_continuous(breaks=seq(-3,3,0.5), limits=c(-3,3))+
  scale_y_discrete(labels = c("2012","2013","2014","2015","2016"))+
  scale_fill_viridis(end=0.8,discrete=T)+
  scale_color_viridis(end=0.8,discrete=T)+
  guides(fill=FALSE, color=FALSE)+
  xlab("Parameter Value")+
  ylab("Year")+
  theme_bw()+
  theme(panel.grid.minor = element_blank(),
        panel.border     = element_blank(),
        axis.line        = element_blank(),
        axis.ticks       = element_blank())

\newpage{}

### 2017-08-15 manuscript/usses_water_ms.Rmd
updates all model related stuff to ARK specification

## Data Analysis
Our main goal was to test whether the relationship between ANPP and soil moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to simply as `soil moisture').
To achieve this goal, we fit a multi-level random intercept and random slope regression with log(ANPP) as the response variable and soil moisture as the sole predictor.
We log-transformed ANPP to account for heteroscedasticity.
We fit the model under a Bayesian framework, allowing us to test for treatment differences by comparing the posterior distributions of the treatment-level coefficients [e.g., @Tredennick2013].
Both log(ANPP) and soil moisture were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].

@@

\begin{align}
\mu_{j(k(t))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_{k(t)} + \gamma_t, \\
y_{j(k(t))} &\sim \text{Normal} \left(\mu_{j(k(t))}, \sigma^2_{k} \right),
\end{align}

\noindent{}where $\mu_{j(k(t))}$ is the deterministic prediction from the regression model for plot *j* associated with treatment *k* in year *t*, $\beta_{0,j(k)}$ is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
Data include the standardized log(ANPP) observations $\left( y_{j(k(t))} \right)$ and soil moisture ($x_{k(t)}$).

@@

\noindent{}where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\boldsymbol{\Sigma}(k)$).
The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
The random year effects ($\boldsymbol{\gamma}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a weibull distribution.
A full description of model is in Appendix 2.

@@

\begin{description}
\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
\item [Appendix 2.] Details of the hierarchical Bayesian model and Fig. A2-1.
\item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
\end{description}

@@

```{r table, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
library(xtable)
coef_probs <- readRDS("../results/betadiff_probabilities.RDS")
coef_caption <- "Probability that the difference between treatment-level coefficients is greater or less than zero for each of the precipitation treatments. Probabilities denote max(Pr(difference) $>$ 0, Pr(difference) $<$ 0), where the sign of the coefficient was ignored. Thus, probabilities represent the probability of the compared coefficients being different from one another. Pr(difference) values far from 0 or 1 indicate no support for the difference being greater than zero. Bold coefficients denote Pr(difference) $\\geq$ 0.95 (there are none)."
print(xtable(coef_probs, caption = coef_caption, digits = 2),
      caption.placement="top",
      include.rownames = F,
      sanitize.colnames.function = identity,
      comment = FALSE,
      size = "normalsize")

@@

# Make the table
inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis` tests whether treatments form unique clusters in multidimensial space; `betadisper` tests whether treatments have similar dispersion. For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
print(xtable(dplyr::select(sppcomp_stats, -Sums, -Means), caption = inyear_caption, digits = 2),
       caption.placement="top",
       include.rownames = F,
       sanitize.colnames.function = identity,
       comment = FALSE,
       size = "normalsize")
```

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/glmm_main_results.png}
  \caption{Results from the generalized linear mixed effects model. (A) Posterior distributions of treatment-level parameters ('Intercept' and the effect of 'Soil Moisture'). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity. (B) Scatterplot of the data and model estimates shown a solid lines. Model estimates come from treatment level coefficients (colored lines) and the overall mean coefficients (heavy black line). Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig. A2-1.}
\end{figure}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/glmm_yeardiffs.png}
  \caption{Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
\end{figure}

@@

<!---
\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{../figures/rank_clocks_scaled_abundance.png}
  \caption{Rank clocks of average species' scaled abundance by treatment. (A) All species. (B) Dominant species.}
\end{figure}
---->

>>>>

## Data Analysis
Our main goal was to test whether the relationship between ANPP and soil moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to simply as `soil moisture').
To achieve this goal, we fit a hierarchical regression model with log(ANPP) as the response variable and soil moisture as the sole predictor.
We log-transformed ANPP to account for heteroscedasticity.
Both log(ANPP) and soil moisture were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].

@@

<!---
\begin{align}
\mu_{j(k(t))} &= \beta_{0,j(k)} + \beta_{1,j(k)}x_{k(t)} + \gamma_t, \\
y_{j(k(t))} &\sim \text{Normal} \left(\mu_{j(k(t))}, \sigma^2_{k} \right),
\end{align}
---->

\begin{align}
\mu_{i} &= \boldsymbol{\beta}\textbf{x}_i + \boldsymbol{\gamma}_{j(i)}\textbf{z}_i + \eta_t, \\
\textbf{y} &\sim \text{Normal} \left(\boldsymbol{\mu}, \sigma^2 \right),
\end{align}

\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation *i*, which is associated with plot *j* and treatment year *t*. $\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$. Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of volumetric water content (VWC), binary "drought" value times VWC, and binary "irrigation" value times VWC. Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments. In reference to our model, the hypotheses we wish to test are: (1) the coefficient for drought$\times$VWC is positive and different from zero, and (2) the coefficient for irrigation$\times$VWC is negative and different from zero

To account for the fact that observations within plots and years are not independent, we include three random effects. Specifically, we include plot-specific offsets ($\boldsymbol{\gamma}$) for the intercept and slope terms and year-specific intercept offsets ($\eta_t$). The covariate vector $\textbf{z}_i$ for each observation *i* has two elements: a 1 for the intercept and the scaled value of VWC for that plot and year. The plot-specific coefficients are modeled hierarchically, where plot level coefficients are drawn from a multivariate normal distribution with mean 0 and a variance-covariance strucutre that allows the intercept and slope terms to be correlated:

\begin{align}
\boldsymbol{\gamma}_{j(i)} &\sim \text{MVN} \left( 0, \Sigma  \right),
\end{align}

\noindent{}where $\Sigma$ is the variance-covariance matrix and *j(i)* reads as "plot *j* associated with observation *i*". The random year effects ($\boldsymbol{\eta}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a half-cauchy distribution. A full description of our model is in Appendix 2.

<!----
is the intercept for plot *j* associated with treatment *k*, $\beta_{1,j(k)}$ is the slope term for the effect of soil moisture for plot *j* associated with treatment *k*, $\gamma_t$ is the intercept offset for year *t*, and $\sigma^2_k$ is the process variance for treatment *k*.
Data include the standardized log(ANPP) observations $\left( y_{j(k(t))} \right)$ and soil moisture ($x_{k(t)}$).

@@

\noindent{}where $\boldsymbol{\beta}_{j(k)}$ is the vector of regression coefficients (intercept and slope) for plot *j* associated with treatment *k*, $\boldsymbol{\beta}_{k}$ is the vector of coefficients for each treatment, and $\boldsymbol{\beta}$ is the vector of overall coefficients.
The plot- and treatment-level coefficients are drawn from multivariate normal distributions with covariance matrix $\boldsymbol{\Sigma}$.
For the plot-level coefficients, each treatment has its own variance-covariance matrix (i.e., $\boldsymbol{\Sigma}(k)$).
The overall coefficients are drawn from a normal prior with mean 0 and standard deviation 1.
The random year effects ($\boldsymbol{\gamma}$) are drawn from a normal prior with mean 0 and standard deviation $\sigma_{\text{year}}$, which was drawn from a weibull distribution.
A full description of model is in Appendix 2.
---->

@@

\begin{description}
\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
\item [Appendix 2.] Details of the hierarchical Bayesian model, Fig. A2-1, and Fig. A2-2.
\item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
\end{description}

@@

```{r table, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
library(xtable)
effect_summary <- readRDS("../results/effect_summary.RDS")
colnames(effect_summary) <- c("Coefficient", "Treatment", "Posterior Mean", "Posterior Median",
                              "Lower 95\\% BCI", "Upper 95\\% BCI")
coef_caption <- "Summary statistics from the posterior distributions of coefficients for each treatment."
print(xtable(effect_summary, caption = coef_caption, digits = 2),
      caption.placement="top",
      include.rownames = F,
      sanitize.colnames.function = identity,
      comment = FALSE,
      size = "normalsize")

@@

# Make the table
inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
print(xtable(dplyr::select(sppcomp_stats, -Sums, -Means), caption = inyear_caption, digits = 2),
       caption.placement="top",
       include.rownames = F,
       sanitize.colnames.function = identity,
       comment = FALSE,
       size = "normalsize")
```

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/glmm_main_results.png}
  \caption{Results from the generalized linear mixed effects model. (A) Posterior distributions of the intercept and slope offsets for the drought and irrigation treatments. Offsets indicate the amount to which the coefficients for drought or irrigation treatments differ from the control treatment estimates. Probabilities (``Pr $<$ 0 ='') for each distribution indicate the probability that coefficient is less than zero. Probabilities greater than 0.95 indicate strong support for the coefficient being less than zero. We only show the one-tailed probability for the value being less than zero because the median of each distribution is less than zero. Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity. (B) Scatterplot of the data and model estimates shown a solid lines. Model estimates come from treatment level coefficients (colored lines). Note that we show log(ANPP) on the \emph{y}-axis of panel B; this same plot can be seen on the arithmetic scale in supporting material Fig. A2-1.}
\end{figure}

<!---
\newpage{}

\begin{figure}[!ht]
  \centering
      \includegraphics[width=5in]{../figures/glmm_yeardiffs.png}
  \caption{Posterior distributions of random year effects (intercept offsets). Kernel bandwidths of posterior densities were adjusted by a factor of 5 for visual clarity.}
\end{figure}
--->

@@
